---
title: "H2O Workshop"
subtitle: Airlines Dataframe
output:
  html_document:
    df_print: paged
---

## Set your environment

```{r environment}
# Install snippet
# source('install.R')

setwd(getwd())

library(h2o)
h2o.init(nthreads=-1, max_mem_size="2G")
h2o.removeAll() ## clean slate - just in case the cluster was already running
h2o.clusterStatus()
```

## Data wrangling

[Data information](http://stat-computing.org/dataexpo/2009/the-data.html)

Prepare your data for modelling.

```{r data}
flights.hex <- h2o.importFile(path = 'data/allyears2k.csv', 
                     destination_frame = 'flights.hex')
flights.hex[1:5, ]

flights.hex["Year"] <- as.factor(flights.hex["Year"])
flights.hex["Month"] <- as.factor(flights.hex["Month"])
flights.hex["DayOfWeek"] <- as.factor(flights.hex["DayOfWeek"])
flights.hex["Cancelled"] <- as.factor(flights.hex["Cancelled"])
flights.hex['FlightNum'] <- as.factor(flights.hex['FlightNum'])


h2o.summary(flights.hex, exact_quantiles=TRUE)
```

Response is _IsArrDelayed_ and is a numerical column. 

### Data exploration

Quantiles and histograms

- Quantile: A sorted sample is divided into equal-sized, adjacent, subgroups. The quartiles are the 25, 50, and 75 quantiles. The percentiles are quantiles that divide a distribution into 100 equal parts. The median is the 50 quantile.

- Histogram: A function that counts the number of observations that fall into each of the disjoint categories (bins).

```{r quantiles}
quantile(x =  flights.hex$ArrDelay, na.rm = TRUE)
h2o.hist(flights.hex$ArrDelay)
```


Cancellations ratio

```{r cancelations}
cancellations_month <- h2o.group_by(data = flights.hex, by = "Month", 
                                    sum("Cancelled"), 
                                    gb.control= list(na.methods="rm"))
cancellations_month

by_month <- h2o.group_by(flights.hex, by = "Month", nrow("Month"), 
                         gb.control=list(na.methods="rm"))

cancellation_rate <- cancellations_month$sum_Cancelled/by_month$nrow
rates_table <- h2o.cbind(by_month$Month, cancellation_rate)
rates_tableDF = as.data.frame(rates_table)
plot(rates_tableDF, type = 'h')
```
```{r is_delay}
delayed <- h2o.group_by(data = flights.hex, by = "IsArrDelayed", 
                                    nrow("IsArrDelayed"), 
                                    gb.control= list(na.methods="rm"))
delayed
```

## Modeling

### Set response

Set predictor and response variables

```{r predictor}
response <- "IsArrDelayed"
predictors <- c("Year", "Month", "DayofMonth", "DayOfWeek", "UniqueCarrier", 
      "Origin", "Dest","Distance")
```

Should we include _"IsDepDelayed"_?

### Split the data

We are going to build models with two splittings, one with validation data and one without it.

```{r split}
data_validation <- h2o.splitFrame(flights.hex, ratios=c(.7,.15),
                      destination_frames = c("train","test","valid"),
                      seed = 511892)
names(data_validation) <- c("Train","Test","Valid")

data <- h2o.splitFrame(flights.hex, ratios=c(.85),
                      destination_frames = c("train","test"),
                      seed = 511892)
names(data) <- c("Train","Test")
```

### Model

```{r model}
flights_glm <- h2o.glm(model_id = "unregularized", training_frame=data_validation$Train, 
                       validation_frame = data_validation$Valid, x = predictors, 
                       y = response, family = 'binomial', seed = 511892)
# Each row of the matrix represents the instances in a predicted class while each column represents the instances in an actual class
h2o.confusionMatrix(flights_glm, valid=TRUE)
```
```{r roc_curve}
# false positive rate
fpr <- flights_glm@model$training_metrics@metrics$thresholds_and_metric_scores$fpr
# true positive rate 
tpr <- flights_glm@model$training_metrics@metrics$thresholds_and_metric_scores$tpr
fpr_val <- flights_glm@model$validation_metrics@metrics$thresholds_and_metric_scores$fpr
tpr_val <- flights_glm@model$validation_metrics@metrics$thresholds_and_metric_scores$tpr
plot(fpr,tpr, type='l', main = 'AUC')
lines(fpr_val,tpr_val,type='l',col='red')
legend("bottomright",c("Train", "Validation"),col=c("black","red"),lty=c(1,1),lwd=c(3,3))
```

```{r model_cross-validation}
flights_cross <- h2o.glm(model_id = "cross-validation", training_frame = data$Train, 
                         x = predictors, y = response, family='binomial', 
                         nfolds = 5, seed = 511892)

h2o.confusionMatrix(flights_cross, valid=FALSE) # get confusion matrix in the training data
```

```{r model_regularized}
flights_glm_r <- h2o.glm(model_id = "regularized", training_frame = data$Train, 
                         validation_frame = data$Valid, x = predictors, y = response,
                         family='binomial', lambda_search = TRUE, 
                         solver = 'COORDINATE_DESCENT', alpha = 0.5)
```

```{r explore_model}
flights_glm_r
```


```{r}
h2o.auc(flights_glm_r,valid=FALSE) # on train                   
flights_glm_r@model$training_metrics@metrics$max_criteria_and_metric_scores   
```

```{r predict}
pred <- h2o.predict(object = flights_glm_r, newdata = data$Test)
plot(h2o.performance(flights_glm_r), type = "roc")
```