---
title: "H2O Workshop"
output: html_notebook
---

## Requirements

For compatibility with H2O:

1. Operating Systems:

  - Windows 7 or later
  - OS X 10.9 or later
  - Ubuntu 12.04
  - RHEL/CentOS 6 or later

2. Languages: 
  - Scala (2.10 or later), R (3 or later), and Python (2.7.x, 3.5.x, 3.6.x) are not required to use H2O unless you want to use H2O in those environments.
  - Java is always required. Supported versions include:
    - Java 7 or later. Note: Java 9 is released, but is not currently supported. 
  - To build H2O or run H2O tests, the 64-bit JDK is required.
  - To run the H2O binary using either the command line, R, or Python packages, only 64-bit JRE is required.

3. Browser: An internet browser is required to use H2O’s web UI, Flow. 
  - Latest version of Chrome, Firefox, Safari, or Internet Explorer.

### About H2O

H2O is fast, scalable, open-source machine learning and deep learning. Advanced algorithms such as deep learning, boosting, and bagging ensembles are built-in.

[Download H2O](https://www.h2o.ai/h2o/) to do some machine learning.

[Documentation](http://docs.h2o.ai/h2o/latest-stable/h2o-docs/index.html).

## Set your environment

```{r environment}
# Install snippet
# source('install.R')

library(h2o)
h2o.init()
# h2o.clusterInfo()
```

## Data wrangling

Prepare your data for modelling.

```{r data}
# Dataset
# https://archive.ics.uci.edu/ml/datasets/Occupancy+Detection+

df <- h2o.importFile(path = 'occupancy_data/data.txt', 
                     destination_frame = 'occupancy.hex',
                     col.types = c("time", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric"))
df[1:5, ]
summary(df, exact_quantiles=TRUE)
```

Quantiles and histograms

```{r quantiles}
quantile(x = df$Occupancy, na.rm = TRUE)
h2o.hist(airlines.hex$ArrDelay)
```